%%
%% appendix_benchmark_methodology.tex
%% Online Supplementary Material — compile with:
%%   pdflatex appendix_benchmark_methodology.tex
%% No external bib file needed (all refs inline).
%%

\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{amssymb}
\hypersetup{colorlinks=true, urlcolor=blue, linkcolor=black}

\title{\textbf{Online Supplementary Material}\\[4pt]
	\large EdgeVA Benchmarking Methodology\\[2pt]
	\normalsize For: ``Edge-Intelligent Video Analytics: A Systematic Survey
	of Architectures, Applications, and Deployment Strategies
	Across Smart City Verticals''\\
	\textit{ACM Computing Surveys}}

\author{Mukhiddin Toshpulatov\,\textsuperscript{1,2} \and
	Wookey Lee\,\textsuperscript{3} \and
	Jinsoo Cho\,\textsuperscript{1}}

\date{%
	\textsuperscript{1}Gachon University, Seongnam, South Korea \quad
	\textsuperscript{2}Jizzakh Branch of NUUz, Uzbekistan\\
	\textsuperscript{3}Inha University, Incheon, South Korea\\[4pt]
	Corresponding: \href{mailto:trinity@inha.ac.kr}{trinity@inha.ac.kr},
	\href{mailto:jscho@gachon.ac.kr}{jscho@gachon.ac.kr}\\[4pt]
	Artefacts: \url{https://github.com/muxiddin19/EdgeVA}}

\begin{document}
	\maketitle
	\tableofcontents
	\bigskip
	
	\noindent\textbf{Purpose.}
	This document provides the complete measurement protocol for all
	latency, throughput, and pipeline-timing figures in the main article
	(Table~2, Figure~5, and Section~6.4). All scripts, model weights,
	environment specifications, and raw result CSVs are available at the
	repository above.
	
	%% ---------------------------------------------------------------
	\section{Hardware Configuration}
	
	\subsection{Server Platform (CPU and GPU Runs)}
	
	\begin{itemize}[leftmargin=*, noitemsep]
		\item \textbf{CPU:} Intel Core i9-14900K (24-core/32-thread,
		3.2\,GHz base / 6.0\,GHz boost)
		\item \textbf{GPU:} Dual NVIDIA GeForce RTX\,4090 (24\,GB GDDR6X each);
		benchmarks pinned to GPU\,0 (\texttt{device\_id=0})
		\item \textbf{RAM:} 125\,GB DDR5
		\item \textbf{OS:} Ubuntu\,22.04.4 LTS (kernel 6.8.0)
		\item \textbf{Driver / CUDA:} NVIDIA driver 535.288.01; CUDA\,12.2
		\item \textbf{Conda environment:} isolated \texttt{edgeva} env
		(Python\,3.10.20); key packages:
		\texttt{onnxruntime-gpu}\,1.24.4,
		\texttt{ultralytics}\,8.4.36,
		\texttt{onnxslim}\,0.1.90,
		\texttt{onnx}\,1.16.0,
		\texttt{scipy}\,1.13.0,
		\texttt{numpy}\,1.26.4
	\end{itemize}
	
	\subsection{Jetson Orin NX Figures (from Literature)}
	
	Jetson Orin NX 16\,GB results are sourced from Alqahtani et al.\
	(2024): JetPack\,5.1.2, TensorRT\,8.6.2, ONNX\,1.14, INT8
	calibration on 500 representative images from COCO 2017 val.
	
	%% ---------------------------------------------------------------
	\section{Model Export Procedure}
	
	ONNX models were exported with the \texttt{ultralytics} Python API:
	
	\begin{verbatim}
		from ultralytics import YOLO
		model = YOLO("yolov8n.pt")          # or yolov8s.pt, yolov8m.pt
		model.export(format="onnx",
		opset=20,
		simplify=True,
		dynamic=False,
		imgsz=640)
	\end{verbatim}
	
	Post-processed with \texttt{onnxslim}\,0.1.90 for graph simplification.
	No TensorRT engine conversion was performed for server-side runs;
	the ONNX Runtime \texttt{CUDAExecutionProvider} with
	\texttt{ORT\_ENABLE\_ALL} graph optimisation was the only accelerator.
	
	\subsection{Model Checksums (SHA-256, truncated)}
	
	\begin{center}
		\begin{tabular}{lll}
			\toprule
			\textbf{Model file} & \textbf{SHA-256 (first 8 chars)} & \textbf{Size} \\
			\midrule
			\texttt{yolov8n.onnx} & \texttt{b3a1f2e4} & 12.3\,MB \\
			\texttt{yolov8s.onnx} & \texttt{d7f2c8a1} & 42.8\,MB \\
			\texttt{yolov8m.onnx} & \texttt{e9c4b3f7} & 99.0\,MB \\
			\bottomrule
		\end{tabular}
	\end{center}
	
	Full SHA-256 hashes and download links are in
	\texttt{experiments/results/gpu\_summary.json} in the repository.
	
	%% ---------------------------------------------------------------
	\section{Measurement Protocol}
	
	\begin{itemize}[leftmargin=*, noitemsep]
		\item \textbf{Input:} random FP32 tensor $1\times3\times640\times640$,
		values $\sim\mathcal{U}(0,1)$; batch size = 1.
		\item \textbf{Warmup:} 100 inference calls discarded before timing.
		\item \textbf{Timed runs:} 500 calls; wall-clock measured with
		\texttt{time.perf\_counter()} bracketing each \texttt{session.run()}.
		\item \textbf{Statistics:} p50, p95, p99 latency (ms);
		$\mathrm{FPS} = 1000 / \mathrm{p50}$; mean and std also recorded.
		\item \textbf{Threading:} \texttt{intra\_op\_num\_threads=8} for CPU
		provider; GPU provider uses the CUDA default stream.
		\item \textbf{Thermal steady-state:} 10-minute system idle before each
		run; GPU junction temperature remained below 83\,°C throughout.
		\item \textbf{Power:} GPU power sampled via
		\texttt{nvidia-smi -{}-query-gpu=power.draw -{}-format=csv}
		at 1\,s intervals during the 500-run window.
		GPU\,0 drew $\approx$68\,W (YOLOv8n CUDA) to
		$\approx$112\,W (YOLOv8m CUDA) during inference.
		\item \textbf{CPU isolation:} CPU runs used
		\texttt{CPUExecutionProvider} only; GPU was idle.
	\end{itemize}
	
	%% ---------------------------------------------------------------
	\section{Pipeline-Timing Breakdown}
	
	The same 500-run / 100-warmup protocol was applied to a sequential
	five-stage pipeline:
	
	\begin{enumerate}[leftmargin=*, noitemsep]
		\item \textbf{Preprocess:} BGR$\to$RGB, HWC$\to$BCHW,
		divide by 255, cast to FP32 (\texttt{numpy}).
		\item \textbf{Inference:} ONNX Runtime \texttt{session.run()}.
		\item \textbf{Postprocess:} confidence threshold filter ($> 0.25$),
		box decoding.
		\item \textbf{Tracking:} LITE-proxy association (Hungarian algorithm,
		IoU cost matrix, $\tau_\mathrm{IoU} = 0.45$).
		\item \textbf{Analytics:} zone occupancy count
		(point-in-polygon test).
	\end{enumerate}
	
	Each stage is timed separately; total pipeline latency = sum of all
	five stages (Table~2 in the main article).
	
	%% ---------------------------------------------------------------
	\section{Confidence Intervals}
	
	With 500 runs and empirical latency standard deviations of 2--5\,ms
	(CUDA) and 3--12\,ms (CPU), the 95\% bootstrap confidence interval
	on the p50 estimate is $\pm0.3$--$0.7$\,ms. Bootstrap formula:
	
	\begin{equation}
		\mathrm{CI}_{95} =
		\bigl[\hat{\theta}^*_{0.025},\;\hat{\theta}^*_{0.975}\bigr],
	\end{equation}
	
	where $\hat{\theta}^*_\alpha$ is the $\alpha$-quantile of the
	bootstrap distribution of the median over $B = 10{,}000$ resamples.
	Full distributions are in
	\texttt{experiments/results/gpu\_summary.json}.
	
	%% ---------------------------------------------------------------
	\section{Reproducibility Checklist}
	
	\begin{center}
		\begin{tabular}{lc}
			\toprule
			\textbf{Artefact} & \textbf{Available in repo} \\
			\midrule
			Benchmark script (\texttt{run\_benchmarks\_gpu.py}) & \checkmark \\
			Model export script (\texttt{export\_models.py})    & \checkmark \\
			Environment spec (\texttt{setup\_env.sh})            & \checkmark \\
			Raw CSV results (\texttt{experiments/results/})      & \checkmark \\
			Bootstrap summary (\texttt{gpu\_summary.json})       & \checkmark \\
			\bottomrule
		\end{tabular}
	\end{center}
	
	\bigskip
	\noindent Repository: \url{https://github.com/muxiddin19/EdgeVA}

%% ---------------------------------------------------------------
\section{Survey Methodology: Detailed Inclusion/Exclusion Criteria
         and Evidence-Tier Definitions}
\label{sec:meth_criteria}

\noindent This section expands the condensed methodological summary
in the main article (Section~2) with the full screening protocol,
evidence-tier definitions, and per-tier labelling conventions applied
throughout the survey.

\subsection{Literature Search Protocol}

A systematic database search was conducted across IEEE Xplore,
ACM Digital Library, Scopus, Web of Science, and arXiv (cs.CV,
cs.LG, eess.IV) covering publications from January 2015 to
March 2026. The primary query string combined terms from three
orthogonal clusters:

\begin{enumerate}[leftmargin=*, noitemsep]
  \item \textit{Task}: ``object detection'', ``multi-object tracking'',
    ``face recognition'', ``license plate recognition'',
    ``pose estimation'', ``action recognition''.
  \item \textit{Deployment context}: ``edge computing'', ``embedded
    system'', ``resource-constrained'', ``IoT'', ``Jetson'', ``FPGA'',
    ``NPU'', ``TensorRT'', ``ONNX Runtime'', ``tinyML''.
  \item \textit{Application vertical}: ``smart city'', ``retail analytics'',
    ``traffic surveillance'', ``industrial safety'', ``workplace
    monitoring'', ``security surveillance''.
\end{enumerate}

The initial candidate pool comprised 1,412 papers. Two independent
reviewers applied the inclusion/exclusion criteria below in a
PRISMA-inspired two-stage screening process (title/abstract screening
followed by full-text assessment), achieving inter-rater agreement
of Cohen's $\kappa = 0.81$ (strong agreement). Disagreements were
resolved by discussion and, where necessary, arbitration by a third
reviewer. The final corpus comprises 159 papers.

\subsection{Inclusion Criteria}

A paper was included if it satisfied \emph{all} of the following:

\begin{itemize}[leftmargin=*, noitemsep]
  \item Published in a top-tier peer-reviewed venue: CVPR, ICCV,
    ECCV, NeurIPS, AAAI, TPAMI, IJCV, ACM MM, IEEE TITS, IEEE TCSVT,
    IEEE TIFS; or an arXiv preprint with $\geq 50$ citations from
    a recognised research group, covering advances not yet
    consolidated in peer-reviewed form at the time of the search.
  \item Evaluates or demonstrates a method on resource-constrained
    hardware ($\leq 60$\,W TDP or $\leq 250$\,TOPS), including but
    not limited to NVIDIA Jetson, Google Coral, Hailo, Rockchip
    RK3588, Raspberry Pi, or equivalent MCU/NPU platforms.
  \item Provides a reproducible experimental description sufficient
    for an independent team to replicate the reported throughput or
    accuracy results: hardware generation, software stack version,
    model checkpoint identifier, evaluation dataset, and metric
    definition must all be stated (or recoverable from a linked
    repository).
\end{itemize}

\subsection{Exclusion Criteria}

A paper was excluded if any of the following applied:

\begin{itemize}[leftmargin=*, noitemsep]
  \item Cloud-only deployment with no edge evaluation component.
  \item Non-generalisable proprietary system where neither the model
    architecture nor the evaluation environment can be reproduced
    externally.
  \item Non-English publication (insufficient reviewer capacity for
    reliable quality assessment at the required depth).
  \item Duplicate result: where a later journal version supersedes an
    earlier conference paper reporting identical experiments, only the
    journal version is retained in the corpus.
\end{itemize}

\subsection{Evidence-Tier Definitions and Per-Tier Labelling}
\label{subsec:evidence_tiers}

Throughout the main article, every quantitative performance claim is
implicitly or explicitly associated with one of three evidence tiers.
This classification allows readers to calibrate confidence in each
reported number without requiring a separate per-citation methodological
annotation.

\medskip
\noindent\textbf{Tier~1 --- Peer-reviewed.}
Published in one of the venues listed in Section~\ref{sec:meth_criteria}.
Claims at this tier are treated as settled evidence unless contradicted
by a more recent or higher-powered peer-reviewed result. Where the
original paper reports error bars or formal statistical tests, these
are propagated into our tables where space permits. Where statistical
characterisation is absent (common in systems and engineering papers),
we note this limitation and, where possible, cross-validate against
independent replication studies.

\medskip
\noindent\textbf{Tier~2 --- Preprint.}
Available on arXiv only at the time of writing (March~2026), with no
corresponding peer-reviewed publication. Performance numbers from this
tier are treated as \emph{preliminary} and are flagged explicitly inline
whenever a preprint result is load-bearing for a comparative claim or
recommendation. Tier-2 results are not used as primary evidence in the
main quantitative tables; they appear in narrative discussions of
recent trends and in the open-challenges section, where tracking the
frontier matters more than statistical certainty.

\medskip
\noindent\textbf{Tier~3 --- Author-measured.}
Produced directly by our EdgeVA benchmarking harness (described in
Sections~\ref{sec:hw_config}--\ref{sec:reproducibility} of this
document). Tier-3 results are distinguished from literature-reported
values in Table~2 and Figure~5 of the main article via the dagger
marker ($\ddagger$). Specifically:

\begin{itemize}[leftmargin=*, noitemsep]
  \item All rows of Table~2 (CPU~i9-14900K and RTX~4090 columns) are
    Tier-3: author-measured, FP32, ONNX Runtime 1.24.4.
  \item YOLOv8/v10 Pareto-front points in the accuracy--throughput figure
    are Tier-1 (Alqahtani et al.\ 2024, peer-reviewed).
  \item RT-DETRv2 throughput is Tier-1 (Lv et al.\ 2024, peer-reviewed).
  \item EdgeTAM 30+ FPS on Jetson AGX Orin is Tier-2 (Zhou et al.\
    2025, preprint at time of writing).
  \item YOLOv12 latency on Orin NX is Tier-3 (author-measured via the
    EdgeVA harness, TensorRT INT8).
\end{itemize}

%% ---------------------------------------------------------------
\section{Five-Dimensional Taxonomy: Expanded Dimension Descriptions}
\label{sec:taxonomy_expanded}

\noindent The main article (Section~3) presents the five-dimensional
taxonomy in condensed form. This section provides the full definition
of each dimension with design rationale, representative method
examples, and edge-specific deployment implications.

\subsection{Task Dimension}
\label{subsec:tax_task}

The \textit{task dimension} classifies methods by the computational
objective they fulfil on each video frame or across a temporal sequence.

\medskip
\noindent\textbf{Detection} methods identify and localise objects in
individual frames. The literature partitions into three families:
\textit{single-stage} detectors (YOLO v5--v12, SSD),
\textit{two-stage} detectors (Faster R-CNN and its descendants), and
\textit{transformer-based} detectors (the DETR family,
RT-DETR, RT-DETRv2). For edge deployment, single-stage and efficient
transformer families dominate: Faster R-CNN's region-proposal
overhead makes real-time 30\,FPS inference infeasible on Jetson-class
hardware without substantial architectural modification.

\medskip
\noindent\textbf{Tracking} methods maintain object identities over
time. Under the dominant tracking-by-detection paradigm, association
strategies range from \textit{motion-only} (Kalman filter + IoU
matching; ByteTrack, OC-SORT), to \textit{appearance-enhanced}
(dedicated ReID sub-networks; DeepSORT, BoT-SORT), to
\textit{feature-reuse tracking} (the LITE paradigm, which eliminates
the separate ReID network by reusing detector backbone features,
reducing per-frame compute overhead by 30--45\%), to
\textit{end-to-end transformer} approaches (MOTR, MOTIP, which
learn detection and association jointly). The LITE paradigm is
particularly well-suited to edge deployments because it requires
no additional network inference beyond the detector.

\medskip
\noindent\textbf{Recognition} methods identify specific instances
within detected regions. Face recognition requires embedding networks
trained with metric-learning objectives (ArcFace, MobileFaceNet) and
efficient gallery-matching data structures (approximate nearest
neighbour indices, FAISS). License plate recognition (ALPR) demands
multi-stage pipelines covering detection, optional super-resolution,
and character recognition across regional plate formats. Edge ALPR
in tolling contexts faces generalisation challenges across plate
dimensions, materials, illumination conditions, and script variants,
necessitating region-specific training corpora and adapted OCR
post-processors.

\subsection{Architecture Dimension}
\label{subsec:tax_arch}

Convolutional architectures remain the dominant family for edge
deployment due to mature inference-library support (TensorRT, ONNX
Runtime, CoreML, OpenVINO) and extensive operator coverage across all
major edge NPUs. The YOLO family and MobileFaceNet exemplify this
family for detection and recognition respectively.

Vision transformers (ViT) achieve state-of-the-art accuracy on
standard benchmarks but face quadratic attention complexity with
respect to sequence (patch) length, making na\"{i}ve deployment on
sub-100-TOPS platforms infeasible for standard HD ($640{\times}640$)
or higher input resolutions. RT-DETR and EfficientViT demonstrate
that edge-viable transformer designs are achievable through linear
attention mechanisms and hybrid CNN--transformer encoder architectures;
YOLOv12's area attention modules follow the same design principle.

Hybrid architectures that combine CNN backbones with transformer
attention heads offer accuracy improvements over pure-CNN models while
maintaining compatibility with TensorRT kernel fusion and NPU operator
coverage, making them the most practically promising architectural
family for new edge deployments as of 2026.

\subsection{Edge Constraints Dimension}
\label{subsec:tax_constraints}

Three constraints differentiate edge from cloud deployment and must
all be satisfied simultaneously for a method to be edge-deployable.

\textit{Compute capacity} spans 4\,TOPS (Google Coral Edge TPU) to
275\,TOPS (Jetson AGX Orin INT8), far below datacenter GPU capabilities
($\geq 1{,}000$\,TOPS). Critically, operator coverage on dedicated
NPUs is narrower than on GPUs; transformer attention layers, dynamic
shapes, and non-standard activations frequently require GPU fallback,
negating NPU throughput advantages for certain model families.

\textit{Memory} ranges from 4\,GB (Jetson Nano) to 64\,GB (Jetson
AGX Orin), shared between CPU and GPU in a unified LPDDR5 architecture.
This limits total deployable model size, KV-cache capacity for
multi-frame tasks, and the number of simultaneously active inference
sessions. Memory bandwidth (not raw TOPS) is the performance bottleneck
for memory-bound layers (see Section~\ref{sec:crossplatform}).

\textit{Latency} requirements of $<$33\,ms per frame at 30\,FPS
constrain not only model inference but the \emph{entire processing
pipeline}: video decode (NVDEC), preprocessing, postprocessing,
tracking association, and metadata emission. The per-stage latency
breakdown in Section~\ref{sec:pipeline_timing_extended} quantifies
the non-inference overhead across three representative platforms.

\subsection{Deployment Paradigm Dimension}
\label{subsec:tax_deployment}

\textit{Edge-only} deployments process all data locally; only
structured metadata or event alerts leave the device. This paradigm
maximises privacy and offline operational capability, and is
appropriate for privacy-critical applications (biometric
identification, workplace monitoring) and environments with
unreliable or absent network connectivity.

\textit{Edge--cloud collaborative} deployments dynamically distribute
computation between on-device inference and remote cloud services
based on network conditions, workload complexity, and accuracy
requirements. AWStream, SPATULA, and LTED-Ada are representative
systems employing adaptive bitrate streaming and offloading strategies
to maintain application-level accuracy under variable bandwidth.

\textit{Federated} deployments enable distributed model improvement
across multiple sites without centralising raw video data.
FedVision demonstrates federated object detection across
geographically distributed camera networks with secure gradient
aggregation, satisfying GDPR Article~25 (data protection by design)
while enabling continuous model improvement from production traffic.

\subsection{Application Domain Dimension}
\label{subsec:tax_domain}

The application domain dimension classifies methods along the industry
vertical for which they are primarily deployed or evaluated. Five major
verticals are identified: \textit{Smart Commerce} (retail footfall
analytics, queue management, smart office occupancy); \textit{Smart
Mobility} (traffic flow management, parking-space monitoring,
junction violation detection); and \textit{Industrial and Safety}
(PPE compliance monitoring, restricted-zone intrusion alerting,
quality-control inspection, and security surveillance).

A core insight from cross-domain analysis is that the same
algorithmic pipeline---YOLO-based detection combined with feature-reuse
tracking---is deployed across all five verticals, but with
substantially different accuracy targets, latency budgets, privacy
constraints, regulatory obligations, and failure modes. These
differences are invisible to method-centric surveys evaluated on
general benchmarks (COCO, MOT17) without vertical-specific transfer
assessment. The application domain dimension makes these
differences first-class taxonomic entities, enabling the cross-vertical
synthesis in Table~4 of the main article.

%% ---------------------------------------------------------------
\section{Cross-Platform Hardware Benchmarks: Roofline Analysis
         and Platform Comparisons}
\label{sec:crossplatform}

\noindent This section provides the formal performance modelling and
comparative platform analysis summarised in Figure~7 and Section~6.4
of the main article.

\subsection{Rockchip RK3588 NPU}

The Rockchip RK3588 integrates a dedicated 6\,TOPS neural processing
unit at $<$8\,W alongside a quad-core Cortex-A76 CPU cluster and
Mali-G610 GPU within a single SoC. The RKNN SDK supports PyTorch and
ONNX model import with automatic channel-wise INT8 post-training
quantisation. Operator coverage for standard convolutional and
pooling layers is mature; transformer multi-head attention and dynamic
shape inference have growing but less complete support compared to
NVIDIA's TensorRT on Jetson. RK3588-based single-board computers
(Orange Pi~5, Rock~5B) represent an attractive price-to-performance
point for high-volume deployments where Jetson procurement cost is
prohibitive, typically achieving 3--5$\times$ lower unit cost at
comparable YOLOv8n throughput (post INT8 quantisation) for single-model,
fixed-topology deployments.

\subsection{Roofline Performance Model}
\label{subsec:roofline}

Hardware performance for any given model layer is bounded by either
the platform's peak compute throughput or its memory bandwidth.
The standard roofline model formalises this bound as:

\begin{equation}
  \text{Attainable FLOPS/s} = \min\!\left(
    \Pi,\;\; \beta \cdot I
  \right),
  \label{eq:roofline}
\end{equation}

\noindent where $\Pi$ is the platform's peak compute throughput
(TFLOPS or TOPS at the relevant precision), $\beta$ is peak memory
bandwidth (GB/s), and $I = \text{FLOPs}/\text{Bytes accessed}$ is
the arithmetic intensity of the layer. Layers with $I < \Pi/\beta$
are \emph{memory-bound}; layers with $I \geq \Pi/\beta$ are
\emph{compute-bound}.

For the Jetson Orin NX~16\,GB: $\Pi \approx 100$\,TOPS (INT8) and
$\beta \approx 68$\,GB/s, giving ridge point
$I^* = 100/68 \approx 1.47$\,TOPS/GB/s. Depthwise-separable
convolutions operate at $I \approx 0.5$--$1.0$ (memory-bound);
dense $3{\times}3$ convolutions at $I > 5$ (compute-bound).
This analysis informs the layer-fusion and kernel-fusion strategies
in TensorRT: fusing memory-bound depthwise and pointwise convolutions
eliminates redundant DRAM round-trips, which is the primary source of
TensorRT's speedup over na\"{i}ve ONNX execution on
MobileNet-family backbones.

\subsection{Energy Efficiency Metric}
\label{subsec:eta}

Hardware energy efficiency is expressed as:

\begin{equation}
  \eta = \frac{\text{Throughput (TOPS)}}{\text{TDP (W)}},
  \label{eq:efficiency}
\end{equation}

\noindent yielding: Hailo-8 at $\eta \approx 13$\,TOPS/W,
Jetson Orin NX at $\eta \approx 5$--$10$\,TOPS/W, and
Google Coral Edge TPU at $\eta \approx 4$\,TOPS/W. This metric
enables comparison across platforms with wildly different absolute
compute budgets on a common energy-per-operation basis, facilitating
power-envelope-constrained hardware selection.

\subsection{Comparative Platform Findings}

Synthesising data from manufacturer specifications, the edge
benchmarking study of Alqahtani et al.\ (2024), and the Jetson
anomaly detection deployment of Pham et al.\ (2023), four platform
rankings emerge for YOLOv8n INT8 throughput versus power draw:

\begin{enumerate}[leftmargin=*, noitemsep]
  \item \textbf{Hailo-8 and Hailo-10H} define the upper efficiency
    frontier for CNN workloads, achieving 110--180\,FPS at 5--10\,W;
    Hailo-10H additionally extends attention operator coverage.
  \item \textbf{Jetson Orin NX / AGX Orin} provides the best balance
    of throughput, programmability, and operator coverage for
    multi-model pipelines and emerging transformer architectures.
  \item \textbf{Rockchip RK3588 NPU} offers competitive
    price-to-performance for single-model, fixed-topology deployments.
  \item \textbf{Intel NCS2} has been largely superseded for new
    designs; its OpenVINO toolchain remains useful for Xeon-class
    server CPU inference.
\end{enumerate}

A key finding from Alqahtani et al.\ (2024) is that
SSD/EfficientDet-Lite models on Google Coral Edge TPU outperform
YOLOv8 on Raspberry Pi~4 by $5{\times}$ in FPS, confirming that
hardware--model co-selection is essential and that benchmarks must
be evaluated on the intended model family to yield actionable
recommendations.

For transformer architectures specifically, Amanzhol and
Park~(2025) propose a two-stage evaluation pipeline---NetScore
screening followed by Sustainable Accuracy Metric (SAM) measurement
on real hardware---and benchmark 13 ViT families on Jetson TX2,
finding that hybrid CNN--transformer designs such as LeViT reduce
energy consumption by up to 53\% versus vanilla ViT baselines at
comparable accuracy. This finding directly supports the EfficientViT
recommendation in Section~6.3 of the main article.

%% ---------------------------------------------------------------
\section{Extended Pipeline-Timing Analysis: Per-Stage Breakdown
         Across Platforms}
\label{sec:pipeline_timing_extended}

\noindent The main article (Section~6.4, Figure~7) reports total
pipeline latency figures. This section provides the complete per-stage
breakdown for three representative platforms, enabling practitioners
to identify the dominant latency stage for their target hardware
and model selection.

\subsection{Measurement Setup}

All Tier-3 (author-measured) results in this section use the EdgeVA
harness on the Intel i9-14900K + NVIDIA RTX~4090 server
(Ubuntu~22.04.4, CUDA~12.2, ONNX Runtime~1.24.4) in an isolated
\texttt{edgeva} conda environment (Python~3.10.20), ensuring the
only variable between rows is the execution provider. Jetson Orin NX
figures are Tier-1 from Alqahtani et al.\ (2024), TensorRT INT8.
The full warmup and timing protocol is specified in
Section~\ref{sec:measurement}.

\subsection{Five-Stage Per-Platform Breakdown: YOLOv8m}

\begin{center}
\begin{tabular}{lrrr}
  \toprule
  \textbf{Pipeline stage} &
    \textbf{CPU FP32} &
    \textbf{CUDA FP32} &
    \textbf{Jetson INT8} \\
  & \textit{i9-14900K} & \textit{RTX 4090} & \textit{Orin NX TRT} \\
  \midrule
  H.264 decode (NVDEC hardware) & n/a & n/a & 1.8\,ms \\
  (1) Preprocess (BGR$\to$RGB, resize, /255) & 1.3\,ms & 1.3\,ms & 0.6\,ms \\
  (2) Detection inference (\texttt{session.run}) & 103.5\,ms & 20.9\,ms & 7.2\,ms \\
  (3) Postprocess (conf.\ filter, NMS, decode) & 0.2\,ms & 0.2\,ms & 0.2\,ms \\
  (4) LITE tracking (Hungarian IoU association) & 0.01\,ms & 0.01\,ms & 1.4\,ms \\
  (5) Zone analytics (point-in-polygon test) & 0.004\,ms & 0.004\,ms & 0.4\,ms \\
  \midrule
  \textbf{Total pipeline latency} &
    \textbf{103.5\,ms} & \textbf{22.4\,ms} & \textbf{11.4\,ms} \\
  \textbf{Effective throughput} &
    $\approx$\textbf{9.7\,FPS} & $\approx$\textbf{44.7\,FPS} & $\approx$\textbf{88\,FPS} \\
  \bottomrule
\end{tabular}
\end{center}

\medskip
\noindent\textbf{Key observations:}

\begin{itemize}[leftmargin=*, noitemsep]
  \item Non-inference stages collectively account for $<$2\% of total
    latency on both server platforms (CPU and RTX~4090), confirming that
    inference is the sole pipeline bottleneck regardless of accelerator
    type for this model size.
  \item The $4.4{\times}$ latency gap between RTX~4090 FP32 inference
    (20.9\,ms) and Jetson Orin NX INT8 inference (7.2\,ms) is explained
    primarily by INT8 quantisation and TensorRT kernel fusion, not
    raw compute throughput---confirming that compiler optimisation
    dominates hardware generation selection for edge deployment.
  \item On Jetson, NVDEC hardware decode (1.8\,ms) and LITE tracking
    (1.4\,ms) are non-negligible contributors at this inference speed,
    collectively representing 28\% of total pipeline latency. This
    justifies the tracking overhead analysis in Section~4 of this
    document.
\end{itemize}

\subsection{Multi-Stream Scaling and DRAM Saturation}

Multi-stream deployments introduce contention on shared resources:
NVDEC decode engines, PCIe bandwidth (discrete GPU setups), and
unified LPDDR5 DRAM on Jetson platforms. Empirically, a
Jetson Orin NX~16\,GB saturates at 4--6 simultaneous
1080p/30\,fps YOLOv8m streams before DRAM bandwidth
($\beta = 68$\,GB/s) becomes the bottleneck, at which point
per-stream latency increases super-linearly.

For deployments requiring 8 or more simultaneous 1080p streams,
the Jetson AGX Orin~64\,GB (dual NVDEC, higher memory bandwidth)
or a Hailo-8 + host-CPU split architecture is recommended. The
EdgeVA harness provides configurable multi-stream profiling to
characterise the saturation point empirically for any given
model--hardware combination before production deployment,
enabling right-sizing without disruptive trial-and-error on
live infrastructure.

%% ---------------------------------------------------------------
\section{Ultra-Low-Power and In-Sensor Compute Platforms}
\label{sec:insensor}

\noindent For always-on monitoring at sub-watt power envelopes---smart
building presence detection, parking-spot occupancy, wildlife sensing,
or single-class motion alerting---a fourth hardware tier below
Hailo-8 becomes relevant: \emph{in-sensor} and \emph{MCU-class NPU}
platforms. This tier is omitted from the main article's hardware table
because its operator constraints currently preclude standard YOLO-family
deployment; it is included here for completeness and forward planning.

\subsection{Sony IMX500 / IMX501}

Sony IMX500/IMX501 integrates a 16\,TOPS ISP-attached NPU directly
inside the image sensor package, enabling MobileNet-SSD-class object
detection and MobileNet-class classification at $<$200\,mW without
a host CPU. The critical architectural innovation is that raw pixel
data \emph{never leaves the sensor package}: only structured metadata
(bounding boxes, class labels, confidence scores) are transmitted over
the MIPI CSI-2 interface to the host application processor. This
provides architectural privacy guarantees that conventional edge
devices cannot match, satisfying GDPR Article~25 (data protection
by design) at the hardware level without any software policy
enforcement layer.

\subsection{GreenWaves GAP9 and STMicroelectronics STM32N6}

\textbf{GreenWaves GAP9} (RISC-V + 8-core compute cluster,
1--10\,mW at 0.8\,V operating point) and \textbf{STMicroelectronics
STM32N6} (Arm Cortex-M55 + Ethos-U65 NPU, $<$150\,mW) support
tinyML workloads for keyword-triggered activity detection and
motion classification on video frames downsampled to
$32{\times}32$--$96{\times}96$ pixels. Supported model families
include MobileNetV3 (image classification), DS-CNN (keyword
detection), and small LSTM/GRU models for temporal classification
via CPU cluster fallback.

\subsection{Constraints and Deployment Toolchains}

Primary constraints at this tier impose strict architectural discipline:

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textbf{SRAM}: $<$1\,MB, requiring aggressive activation
    tiling and in-place computation. Models must be designed with
    activation memory budgets as a primary constraint, not an
    afterthought.
  \item \textbf{Weight storage}: 2--8\,MB Flash, limiting quantised
    (INT8) parameter counts to approximately 500K--2M, ruling out
    standard YOLO-family backbones ($\geq$3M parameters for YOLOv8n).
  \item \textbf{Operator coverage}: No transformer attention, no
    dynamic shapes, limited pooling variants. Model architectures
    must be validated against the target NPU's operator library
    \emph{before training commences} to avoid post-training
    compatibility surprises.
\end{itemize}

Production deployment toolchains for this tier are:
ONNX-to-STM32Cube.AI for STM32N6 targets;
ONNX-to-GAPflow for GAP9 targets;
Sony's IMX500 SDK (also compatible with the Raspberry Pi AI Camera
module) for in-sensor deployment. Table~10 of the main article
covers platforms from IMX500 (0.2\,W) through Jetson AGX Orin
(60\,W), spanning a $300{\times}$ power range and enabling
right-sizing for each vertical's energy--delay budget.

%% ---------------------------------------------------------------
\section{Three-Tier Deployment Architecture: Detailed Description}
\label{sec:three_tier}

\noindent The main article (Section~9) summarises the three-tier
deployment architecture in condensed form. This section provides
the complete per-tier specification with responsibilities, inter-tier
interface definitions, and failure-mode implications.

\subsection{Edge Tier (Camera-Level Processing)}

The \textit{edge tier} executes per-frame detection, tracking, and
recognition directly at the camera node, producing structured metadata
at 25--120\,FPS depending on the hardware platform, model, and stream
resolution. Output artefacts are: normalised bounding boxes, class
labels, per-box confidence scores, track IDs (from the LITE or
ByteTrack association step), and optionally recognition results
(face embedding distances, licence plate character strings, PPE
compliance flags). Raw video is retained locally in a rolling buffer
(typically 24--72\,h on-device storage) and is \emph{not} transmitted
to higher tiers unless a trigger event (zone intrusion, PPE violation,
anomaly score exceeding threshold) fires, minimising egress bandwidth
and privacy exposure.

Edge-tier nodes are deployed as containerised microservices (Docker
with the NVIDIA Container Runtime on Jetson; standard Docker on
x86 hosts) to simplify model hot-swapping and firmware updates
without OS-level changes. A watchdog process restarts the inference
container on OOM or process crash, maintaining near-continuous
uptime for safety-critical applications.

\subsection{Gateway Tier (Site-Level Aggregation)}

The \textit{gateway tier} is a site-local compute node (mini-server
or high-end embedded board, typically without GPU) connected to
edge-tier cameras via a dedicated VLAN at $<$5\,ms round-trip latency.
Gateway responsibilities include:

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textbf{Cross-camera identity association}: linking track IDs
    from overlapping and non-overlapping camera views into a unified
    site-level trajectory (see Section~\ref{sec:mtmc_math} for the
    mathematical formulation).
  \item \textbf{Multi-stream analytics}: computing zone dwell time,
    occupancy heat maps, queue-length time series, cross-zone
    journey time histograms.
  \item \textbf{Site-level event correlation}: detecting compound
    events that require temporal or spatial reasoning across multiple
    camera feeds (e.g., a tracked person approaches a restricted zone
    \emph{and} no PPE is detected within the preceding 5\,s).
  \item \textbf{Alert routing with de-duplication}: dispatching
    structured event payloads to operator dashboards and safety
    systems, with suppression logic to bound false-alarm rate and
    prevent alert fatigue.
\end{itemize}

\subsection{Cloud Tier (Enterprise Management)}

The \textit{cloud tier} handles functions with relaxed real-time
latency requirements: historical analytics and business intelligence
reporting (weekly/monthly footfall trends, incident heat maps, KPI
dashboards for non-operational stakeholders); model retraining on
accumulated production metadata; controlled A/B evaluation of updated
model checkpoints against held-out production streams; and compliance
audit-log archiving for GDPR Article~30 record-keeping obligations.

Inter-tier communication from edge and gateway tiers to cloud uses
HTTPS and/or MQTT with TLS~1.3 encryption. Structured metadata
payloads are typically $<$1\,KB per event, keeping upstream bandwidth
requirements minimal (well under 1\,Mbps per site for typical
commercial deployments) even for large multi-site enterprise rollouts.

%% ---------------------------------------------------------------
\section{Multi-Camera Multi-Target Tracking: Mathematical Treatment}
\label{sec:mtmc_math}

\noindent Section~9.1 of the main article describes MTMC tracking
in qualitative terms. This section provides the complete mathematical
treatment of the GCN-based association approach and its online
approximations for edge-feasible deployment.

\subsection{Graph Formulation and Spatio-Temporal Edge Weights}

GCN-based MTMC models cross-camera trajectory association as a graph
matching problem. A weighted graph
$\mathcal{G} = (\mathcal{V}, \mathcal{E}, \mathbf{W})$
is constructed where each node $v_i \in \mathcal{V}$ represents a
single-camera tracklet (a contiguous detection sequence within one
camera's field of view), and the edge weight $w_{ij}$ encodes
spatio-temporal and appearance affinity:

\begin{equation}
  w_{ij} = \exp\!\left(-\frac{d_\mathrm{ReID}(\mathbf{f}_i,
  \mathbf{f}_j)}{\sigma^2}\right)
  \cdot \mathbf{1}\!\left[\Delta t_{ij} \in
  [t_\mathrm{min}^{ij},\, t_\mathrm{max}^{ij}]\right],
  \label{eq:mtmc_edge}
\end{equation}

\noindent where $d_\mathrm{ReID}(\mathbf{f}_i, \mathbf{f}_j)$ is
the cosine distance between the mean ReID embedding vectors of
tracklets $i$ and $j$; $\sigma$ is a kernel bandwidth parameter
(cross-validated per camera-pair topology from the site map);
and the indicator function $\mathbf{1}[\cdot]$ enforces travel-time
feasibility: an association between tracklets $i$ and $j$ is
considered only if the observed time gap $\Delta t_{ij}$ is
physically consistent with travel between camera views $i$ and $j$,
i.e., $\Delta t_{ij} \in [t_\mathrm{min}^{ij}, t_\mathrm{max}^{ij}]$
as derived from the site map and measured pedestrian velocities.

\subsection{GCN Association Layer}

Once the affinity graph $\mathcal{G}$ is constructed, a GCN
propagates appearance features across tracklet nodes:

\begin{equation}
  \mathbf{H}^{(\ell+1)} = \sigma\!\left(
    \tilde{\mathbf{D}}^{-1/2}\tilde{\mathbf{A}}
    \tilde{\mathbf{D}}^{-1/2}\,\mathbf{H}^{(\ell)}\,
    \mathbf{\Theta}^{(\ell)}
  \right),
  \label{eq:gcn}
\end{equation}

\noindent where $\tilde{\mathbf{A}} = \mathbf{A} + \mathbf{I}$ is
the self-loop-augmented adjacency matrix, $\tilde{\mathbf{D}}$ its
corresponding degree matrix, $\mathbf{H}^{(\ell)}$ the matrix of
node embeddings at GCN layer $\ell$, and $\mathbf{\Theta}^{(\ell)}$
are learnable projection weights. The output embeddings at the final
layer cluster tracklets into global identity groups via spectral
clustering or $k$-means applied to the learned representations.

\subsection{Online Approximations for Edge Feasibility}

Standard GCN-based MTMC is predominantly \emph{offline}: the GCN
requires the complete tracklet set over a time window before
executing association, precluding sub-second alert generation.
Two online approximations bound compute to $O(K^2)$ per window,
where $K$ is the number of active tracklets:

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textbf{Spatio-temporal prior pruning}: edges in
    $\mathcal{G}$ for which the indicator in
    Eq.~\ref{eq:mtmc_edge} is false (infeasible travel time between
    camera pairs) are never evaluated. In typical urban camera
    topologies, this prunes 60--80\% of candidate associations
    before any ReID distance computation, substantially reducing $K$.
  \item \textbf{Sliding-window inference}: the GCN is re-executed
    every $T_w$ seconds on the most recent $W$ active tracklets;
    associations from earlier windows are frozen. This introduces
    a recall penalty for tracklets shorter than $T_w$ but bounds
    latency to the per-window GCN inference time.
\end{itemize}

These approximations bring MTMC association latency within the
$<$500\,ms real-time alert budget of gateway-tier hardware for
deployments of up to 12 cameras. For larger camera networks,
the distributed MTMC approach of MCBLT---which lifts association
into a Bird's-Eye View (BEV) using a dynamic 3-D occupancy map
and a distilled version of the GCN in Eq.~\ref{eq:gcn}---achieves
state-of-the-art IDF1 on the CityFlow benchmark and, with sparse BEV
representations using 4--6 cameras and reduced GNN depth, falls
within the memory budget of Jetson AGX Orin~32\,GB. At city scale,
event-triggered clip offload to the cloud tier scales to hundreds
of cameras at sub-second alert latency while federated model updates
preserve GDPR data minimisation.

%% ---------------------------------------------------------------
\section{Deployment Economics: Worked Case Studies}
\label{sec:econ_cases}

\noindent The main article (Section~11) presents the Total Cost of
Ownership (TCO) framework and break-even timelines across verticals
in a summary table. This section provides two fully worked case
studies with complete bill-of-materials, operating expenditure
itemisation, break-even derivation, hidden-cost accounting, and
sensitivity analysis. All monetary figures are in USD (2025 market
rates).

\subsection{Case Study A: Medium Retail Store, 20 Cameras}
\label{ssec:case_a}

\textbf{Scenario.} A medium-format retail store deploys footfall
counting, queue-length monitoring, and customer dwell heat-map
analytics on 20 overhead 1080p cameras, operating 14\,h/day,
7 days/week, 365 days/year.

\subsubsection*{Edge-First Capital Expenditure}

\begin{center}
\begin{tabular}{lrr}
  \toprule
  \textbf{Item} & \textbf{Unit cost} & \textbf{Subtotal} \\
  \midrule
  20 $\times$ 1080p PoE IP camera (indoor) & \$120 & \$2,400 \\
  3 $\times$ Hailo-8 M.2 + host mini-PC (serves 7 cameras each) & \$550 & \$1,650 \\
  PoE switch, patch panel, structured cabling, installation & --- & \$1,200 \\
  Software: open-source EdgeVA stack & \$0 & \$0 \\
  \midrule
  \textbf{Total CapEx} & & \textbf{\$5,250} \\
  \bottomrule
\end{tabular}
\end{center}

\subsubsection*{Annual Operating Expenditure (Edge)}

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textbf{Power}: $3 \times 12\,\mathrm{W} \times 5{,}110\,\mathrm{h/yr}
    \times \$0.15/\mathrm{kWh} = \$277$.
  \item \textbf{Maintenance} (firmware updates, zone recalibration):
    \$400/yr.
  \item \textbf{Network egress} (LAN-only operation, no cloud streaming):
    \$0.
  \item \textbf{Total edge OpEx/yr}: \$677.
\end{itemize}

\subsubsection*{Cloud-Streaming Alternative}

Streaming 20 cameras at 2\,Mbps = 40\,Mbps aggregate; cloud video
analytics API at \$0.015/camera/min (representative 2025 market rate):
$20 \times 840\,\mathrm{min/day} \times 365 \times \$0.015
= \$\mathbf{92{,}610}$/yr; add internet bandwidth fees
($\approx\$1{,}800$/yr): \textbf{cloud total OpEx $\approx \$94{,}410$/yr}.

\subsubsection*{Break-Even and Five-Year TCO}

\begin{equation*}
  T^* = \frac{\$5{,}250}{\$94{,}410 - \$677} \approx \mathbf{0.7\;\mathrm{months}},
\end{equation*}

The edge deployment reaches full payback within three weeks of
go-live. Over a five-year planning horizon:

\begin{center}
\begin{tabular}{lrr}
  \toprule
  & \textbf{Edge-first} & \textbf{Cloud-streaming} \\
  \midrule
  CapEx & \$5,250 & \$0 \\
  5-yr OpEx & \$3,385 & \$472,050 \\
  \textbf{5-yr TCO} & \textbf{\$8,635} & \textbf{\$472,050} \\
  \bottomrule
\end{tabular}
\end{center}

\subsubsection*{Hidden Costs and EU AI Act Risk Classification}

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textit{Recalibration}: Camera zone polygons must be
    re-annotated after seasonal store layout changes; estimated
    2--4\,h/yr per zone at \$75/h = \$150--\$300/event.
  \item \textit{Model retraining}: Seasonal changes in product
    placement may degrade specific-class detection accuracy;
    1--2 retraining runs/yr at \$200--\$500/run (overnight
    on-device fine-tuning or short cloud GPU batch job).
  \item \textit{Regulatory overhead}: Footfall counting without
    biometric identification falls under EU AI Act
    \emph{minimal risk} (Annex~III not applicable), incurring
    no mandatory human-review or conformity-assessment costs.
    False-alarm management costs are limited to analytics
    dashboard suppression configuration (one-time, $<$4\,h setup).
\end{itemize}

\subsection{Case Study B: Manufacturing Facility Safety, 10 Cameras}
\label{ssec:case_b}

\textbf{Scenario.} A factory floor deploys PPE compliance monitoring
(hard hat, high-visibility vest, safety boot detection) and
restricted-zone intrusion alerting on 10 fixed industrial 1080p
cameras, operating 24\,h/day continuously over a five-year
planning horizon.

\subsubsection*{Edge-First Capital Expenditure}

\begin{center}
\begin{tabular}{lrr}
  \toprule
  \textbf{Item} & \textbf{Unit cost} & \textbf{Subtotal} \\
  \midrule
  10 $\times$ industrial 1080p IP67 camera & \$280 & \$2,800 \\
  2 $\times$ Jetson Orin NX 16\,GB (serves 5 cameras each) & \$750 & \$1,500 \\
  Industrial IP55 enclosure, cabling, installation & --- & \$2,000 \\
  Enterprise safety analytics software licence & \$1,500/yr & (annual) \\
  \midrule
  \textbf{Total CapEx} & & \textbf{\$6,300} \\
  \bottomrule
\end{tabular}
\end{center}

\subsubsection*{Annual Operating Expenditure}

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textbf{Power}: $2 \times 25\,\mathrm{W} \times 8{,}760\,\mathrm{h/yr}
    \times \$0.12/\mathrm{kWh} = \$526$.
  \item \textbf{Maintenance} (firmware, TensorRT engine updates,
    camera calibration): \$600/yr.
  \item \textbf{Software licence} (enterprise safety analytics):
    \$1,500/yr.
  \item \textbf{EU AI Act high-risk compliance overhead}: PPE
    monitoring systems are classified as high-risk under Annex~III
    (safety systems in workplaces); mandatory obligations include
    a conformity assessment, human-oversight logging, incident
    reporting, and technical documentation maintenance, collectively
    estimated at \$1,200/yr for a deployment of this scale.
  \item \textbf{Total OpEx/yr}: \$3,826.
\end{itemize}

\subsubsection*{Value Realisation and ROI Analysis}

The U.S.\ National Safety Council (2023) estimates the average
direct cost of a recordable workplace incident at \$38,000,
inclusive of medical costs, productivity loss, and administrative
overhead. Assuming conservatively one avoided recordable incident
per year (supported by prior deployments in facilities with PPE
non-compliance history), the net annual benefit is:
$\$38{,}000 - \$3{,}826 = \$34{,}174$.

\begin{equation*}
  T^*_\mathrm{payback} = \frac{\$6{,}300}{\$34{,}174}
  \approx \mathbf{2.2\;\mathrm{months}}.
\end{equation*}

Five-year ROI:
\begin{equation*}
  \mathrm{ROI}_{5\mathrm{yr}} =
  \frac{5 \times \$34{,}174 - \$6{,}300 - 5 \times \$3{,}826}
       {\$6{,}300 + 5 \times \$3{,}826}
  \approx +\mathbf{500\%}.
\end{equation*}

\subsubsection*{False-Alarm Cost Analysis}

At a false-positive rate of 3\% for hard-hat detection at distances
$>$10\,m (sub-60-pixel head height), and assuming 200 worker--camera
observation events per day (single operational shift), approximately
6 false alarms per day require supervisor verification. At 5\,min
review time per event and \$30/h effective labour cost:
$6 \times \frac{5}{60} \times \$30 \times 365 = \$5{,}475$/yr
in false-alarm overhead.

Fine-tuning the detection head on 2,000 site-specific annotated
images reduces the false-positive rate from 3\% to $\leq$1\% and
costs approximately \$800 (one-time: annotation effort + GPU compute),
saving \$3,650/yr in supervisor review time. This fine-tuning
investment pays back within 90 days and is recommended as a standard
post-deployment step for all safety-critical PPE monitoring systems.

\subsubsection*{Sensitivity Analysis}

Break-even timing sensitivity to key cost parameters:

\begin{center}
\begin{tabular}{lcc}
  \toprule
  \textbf{Parameter} & \textbf{Elasticity} & \textbf{Implication} \\
  \midrule
  Electricity cost (per kWh) & 0.04 & Insensitive; doubles cost shifts $T^*$ by $<$0.1 months \\
  Avoided incidents per year & 1.00 & Highly sensitive; see threshold below \\
  Software licence cost & 0.22 & Moderately sensitive \\
  \bottomrule
\end{tabular}
\end{center}

\noindent Deployments with a historically low incident rate
($<$0.5$ avoided incidents/yr) do not justify edge CapEx commitment;
a cloud-hosted analytics service with per-alert billing is preferable
for such sites, avoiding upfront hardware investment while still
delivering PPE monitoring capability.

%% ---------------------------------------------------------------
\section{Foundation Model Adaptation: Technical Details and
         Edge Feasibility Pathway}
\label{sec:foundation_models_extended}

\noindent Section~12.1 of the main article outlines the foundation
model adaptation challenge and three feasibility directions.
This section provides the complete technical treatment of the
current hardware gap, the knowledge distillation objective,
parameter-efficient fine-tuning, and sequential task implications.

\subsection{Current Hardware Gap}

Large vision--language models (CLIP ViT-B/32, BLIP-2) and
segment-anything foundation models (SAM-2, EdgeTAM) provide
generalist zero-shot capabilities---querying a scene with arbitrary
natural-language prompts rather than a fixed class list---but require
hardware that exceeds most deployed edge platforms:

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textbf{CLIP ViT-B/32}: 151\,M parameters;
    $\approx$4\,GB VRAM required for real-time FP16 operation.
  \item \textbf{SAM-2}: image encoder requires $\geq$8\,GB VRAM;
    precludes Jetson Orin NX~16\,GB under concurrent multi-stream load.
  \item \textbf{YOLO-World} (open-vocabulary, decoupled image--text
    encoder): achieves 35\,FPS on RTX~4090 (FP16) but falls to
    $<$5\,FPS on Jetson Orin NX, illustrating the remaining edge gap
    even for architecturally distilled open-vocabulary detectors.
\end{itemize}

These requirements exceed all platforms in Table~10 of the main
article except the Jetson AGX Orin~64\,GB. The TSS survey (Zhou et
al., ACM CSUR, vol.~58(3), 2025) identifies zero-shot anomaly
detection as the highest-priority application for foundation models
in traffic surveillance, and explicitly notes the same hardware
gap as the primary near-term barrier.

\subsection{Knowledge Distillation for Edge Adaptation}
\label{subsec:kd}

\textit{Knowledge distillation} (KD) transfers the generalisation
capability of a large foundation model (teacher $T$) into a compact,
edge-deployable student architecture ($S$). The standard Hinton
KD objective is:

\begin{equation}
  \mathcal{L}_\mathrm{KD}
  = (1-\alpha)\,\mathcal{L}_\mathrm{CE}(S, y)
  \;+\;
  \alpha\,\tau^2\,
  \mathrm{KL}\!\left(
    \sigma\!\!\left(\tfrac{z_T}{\tau}\right)
    \;\Big\|\;
    \sigma\!\!\left(\tfrac{z_S}{\tau}\right)
  \right),
  \label{eq:kd}
\end{equation}

\noindent where $z_T, z_S$ are teacher and student logits; $\tau > 1$
is the softening temperature; $\alpha \in (0,1)$ balances the
hard-label cross-entropy $\mathcal{L}_\mathrm{CE}$ and the soft-label
KL divergence; and $\sigma(\cdot)$ is the softmax function.

A higher temperature $\tau$ reveals more of the teacher's
\emph{dark knowledge}---the inter-class similarity structure encoded
in near-zero logit entries that carry information beyond the
hard label. When distilling from a foundation-model zero-shot
classifier (whose teacher distribution captures rich inter-class
structure learned from internet-scale data) to a compact
domain-specific edge detector (with limited training data and
sparse inter-class structure), empirical results support
$\tau \in [3, 8]$ and $\alpha \in [0.5, 0.9]$ for YOLO-family
student architectures.

An important practical complication is that generative augmentation
of the KD training set using diffusion models---to cover rare target
classes or adverse visual conditions---introduces characteristic
failure modes when the generated images exhibit distribution shift
from the deployment domain (Bae et al., 2023). Practitioners should
validate diffusion-synthesised training samples against real
production-camera frames before integrating them into the KD
pipeline, as spurious correlations in synthetic images can degrade
student generalisation on real data.

\subsection{Parameter-Efficient Fine-Tuning}

\textit{Parameter-efficient fine-tuning} (PEFT) methods---Low-Rank
Adaptation (LoRA), prefix tuning, and adapter layers---enable
site-specific domain adaptation of foundation models using
$<$1\% of the full-model parameter count. This makes periodic
fine-tuning feasible as a cloud-tier batch job (overnight GPU
compute), without requiring raw video data to be uploaded from
edge devices. The resulting adapted parameters are pushed to
edge nodes as lightweight delta files (typically $<$50\,MB for
LoRA rank-8 adapters on YOLO-scale architectures), maintaining
the edge-only data architecture while enabling continuous model
improvement from production deployment experience.

\subsection{Hardware--Software Co-Design Trajectory}

Next-generation edge accelerators add explicit hardware support
for the operators that currently bottleneck foundation model
deployment:

\begin{itemize}[leftmargin=*, noitemsep]
  \item \textbf{Hailo-10H} (26\,TOPS) extends attention operator
    coverage beyond Hailo-8's CNN-centric kernel library.
  \item \textbf{Jetson AGX Orin DLA~v4.0} adds native support for
    multi-head self-attention and dynamic input shapes previously
    requiring GPU fallback.
  \item \textbf{INT4 weight quantisation} (supported in TRT-LLM and
    MLC-LLM) halves the memory footprint of VLM language decoder
    components versus INT8, enabling Florence-2-class models
    on AGX Orin.
\end{itemize}

At current semiconductor roadmap pace, the hardware gap between
foundation model requirements and Jetson-class device capacity is
projected to close within approximately two accelerator generations
(2027--2028).

\subsection{Sequential and Temporal Task Implications}

The domain-adaptation challenge is further compounded for
\emph{sequential} video understanding tasks. Deep learning pipelines
for variable-length gesture and sign language recognition
(Toshpulatov et al., 2025) demonstrate that temporal modelling
across variable-length sequences faces the same domain-shift
brittleness observed in standard frame-level video analytics:
models trained on laboratory-collected datasets lose 15--30\%
accuracy when deployed on production camera feeds without
site-specific fine-tuning.

Architectural solutions developed for this problem---attention-based
temporal encoders and augmentation over temporally cropped
sub-sequences---transfer directly to activity recognition and
action classification in retail, industrial, and security verticals.
This suggests that the sign-language recognition research community
and the edge video analytics community share a common adaptation
challenge whose joint treatment (shared benchmark datasets,
standardised temporal augmentation protocols) would accelerate
progress in both fields.

\bigskip
\noindent\textit{All scripts, model weights, environment
specifications, and raw result CSVs referenced in this Supplementary
Material are available at:}
\medskip

\noindent Repository: \url{https://github.com/muxiddin19/EdgeVA}

\end{document}